---
title: "Data analysis example"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

suppose we are giving 2 studens a multiple-choice exam with 40 questions, where each question has 4 choices. We dont know how much the students have studied for this exam, but we think that they will do better than just guessing randomly.

1) what are the parameters of interest?

solution: theta1 = true probability the first student will answer a question correctly, and theta2 = true probability the second student will answer a question correctly.

2) what is our likelihood?

solution: Binomial (40, theta), if we assume that each question is independent and that the probability a student gets each question right is the same for all questions for that student.

3) what prior should we use?

solution: the conjugate prior is a beta prior. Plot the density with dbeta.
```{r}
theta = seq(from=0, to=1, by=.01)
plot(theta, dbeta(theta, 1, 1), type = "l")
plot(theta, dbeta(theta, 4, 2), type = "l")
plot(theta, dbeta(theta, 8, 4), type = "l")
plot(theta, dbeta(theta, 2, 4), type = "l")
```

4) what is the prior probability P(theta > 0.25)? P(theta > 0.5)? P(theta > 0.8)?
```{r}
1 - pbeta(0.25, 8, 4)
1 - pbeta(0.5, 8, 4)
1 - pbeta(0.8, 8, 4)
```

5) suppose the first student gets 33 questions right. What is the posterior distribution for theta1? P(theta > 0.25)? P(theta > 0.5)? P(theta > 0.8)? what is a 95% posterior credible interval for theta1?

solution: Posterior is Beta (8+33, 4+40-33) = Beta(41, 11)
```{r}
# posterior mean
41/(41+11)
# MLE
33/40

# Plot posterior
plot(theta, dbeta(theta, 41, 11), type = "l")

# plot likelihood
plot(theta, dbinom(33, size = 40, p = theta))

# posterior probabilities
1 - pbeta(0.25, 41, 11)
1 - pbeta(0.8, 41, 11)
1 - pbeta(0.5, 41, 11)

# equal-tailed credible interval
qbeta(0.025, 41, 11)
qbeta(0.975, 41, 11)
```

6) suppose the second student gets 24 questions right. What is the posterior distribution for theta2? P(theta > 0.25)? P(theta > 0.5)? P(theta > 0.8)? what is a 95% posterior credible interval for theta2?

solution:

7) what is the posterior probability that theta1 > theta2, i.e., that the first student has a better chance of getting a question right than the second student?

solution:
Estimate by simulation: draw 1000 samples from each and see how often we observe theta1 > theta2
```{r}
theta1 = rbeta(1000, 41, 11)
theta2 = rbeta(1000, 32, 20)
mean(theta1 > theta2)
```

```{r}
plot(theta, dbeta(theta, 5, 1))
```

